KidsAIR: merge exposure datasets

### 2020-04-14 child_health_data_cleaning.Rmd
Update PM data

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 8 April 2020"
output: pdf_document 
---

@@

```{r}
# Load data
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

@@

# Format at-home-log and merge with PM data
```{r}
kids_at_home_log <- read_rds("Output/kids_at_home_log.rds") %>% 
  arrange(area, home_winter_id, at_home_day)

@@

at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day) %>% 
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

#write_rds(at_home_pm_merged, "Output/at_home_pm_merged.rds")
#write_csv(at_home_pm_merged, "Output/at_home_pm_merged.csv")

at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
```

>>>>

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 14 April 2020"
output: pdf_document 
---

@@

```{r}
# Load data
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_linked_ids <- read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
demographics_clean <- 
  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
```

@@

# Format at-home-log and merge with PM data
```{r}
kids_at_home_log <- 
  read_rds(paste0(output_path, "Output/kids_at_home_log.rds")) %>% 
  arrange(area, home_winter_id, at_home_day)

@@

at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

#write_rds(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.rds"))
#write_csv(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.csv"))

at_home_pm_merged <- 
  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
```

### 2020-04-14 dustrak_data_cleaning.Rmd
Update PM data

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)

gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

@@

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "129", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

@@

# Save cleaned dataset
```{r}
kids_pm <- read_rds("Output/kids_pm_new4.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")

kids_pm_new5 <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id)) %>% 
  select(-home, -treatment) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))

# write_rds(kids_pm_new5, "Output/pm_clean.rds")
```

>>>>

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm_new %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)

gaps <<- kids_pm_new %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm_new %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

@@

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_pm_new <- 
  read_rds(paste0(output_path, "Output/kids_pm_new.rds"))

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "273", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

@@

# Check high PM values
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_pm <- 
  read_rds(paste0(output_path, "Output/kids_pm_new4.rds"))

kids_pm_check <- kids_pm %>% 
  filter(pm > 2000) %>% 
  group_by(area) %>% 
  count(home_winter_id)

# 0.1% of the PM data is greater than 2000
# Nothing stood out as being unreasonable when looking (plotting) at homes
# with values over 2000. Plan to exclude values > 10,000, leave the rest
# as is, and do sensitivity analyses excluding home with values over specific
# thresholds. Also plan to see how PM in homes with PM spikes is associated with
# health and behavioral variables.
```

# Join PM data with linked IDs and make new variables for spikes in PM
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_pm <- 
  read_rds(paste0(output_path, "Output/kids_pm_new4.rds")) 
kids_linked_ids <- 
  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))

kids_pm_new5 <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id)) %>% 
  select(-home, -treatment) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
  group_by(home, sampling_day) %>% 
  mutate(pm_spike_1000 = if_else(pm > 1000, "Yes", "No"),
         pm_spike_2000 = if_else(pm > 2000, "Yes", "No"),
         pm_spike_5000 = if_else(pm > 5000, "Yes", "No")) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new)
```

# Save dataset
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

write_rds(kids_pm_new5, paste0(output_path, "Output/pm_clean.rds"))
```

### 2020-04-14 merge_datasets.Rmd
Update PM data

---
title: "KidsAIR: merge exposure datasets"
author: "Ethan Walker"
date: "Started 24 March 2020, Updated 27 March 2020"
output: pdf_document
---

@@

# Load individual datasets
```{r}
#pm_clean <- read_rds("Output/pm_clean.rds")
#sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Create new dataset with mean PM/SUMs; join with covariates
## Long data averaged over sampling date, including SUMs dates, repeated by child_id
```{r}
# Prep PM data to join with SUMs data by sampling day
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ",
           remove = FALSE) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
         pm_daily_obs = n()) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
         pm_total_observations, pm_mean_sampling_period, sampling_day, 
         pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
         sample_date, pm_mean_daily, pm_daily_obs) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

#write_rds(pm_join, "Output/pm_join.rds")
pm_join <- read_rds("Output/pm_join.rds")

@@

#write_rds(sums_join, "Output/sums_join.rds")
sums_join <- read_rds("Output/sums_join.rds")

@@

#write_rds(pm_sums_joined1, "Output/pm_sums_joined1.rds")
pm_sums_joined1 <- read_rds("Output/pm_sums_joined1.rds")

@@

#write_rds(analysis_data7, "Output/exposure_analysis_data_long.rds")
#write_csv(analysis_data7, "Output/exposure_analysis_data_long.csv")

exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")

@@

#write_csv(save_var_names, "Output/kids_data_var_names.csv")
```

# Update long dataset -- REMEMBER TO UPDATE SHORT DATA BELOW
```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
  mutate(sample_date = amb_pm_date) %>% 
  distinct(zip, .keep_all = TRUE)
kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  distinct(zip, .keep_all = TRUE)
wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
  rename_all(tolower)
nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
  rename_all(tolower) 
ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
  rename_all(tolower) 
kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
  mutate(home = gsub(home, pattern = "_", replacement = ""),
         home = gsub(home, pattern = " ", replacement = "")) %>% 
  left_join(kids_amb_pm_data, by = "zip") %>% 
  select(home, zip, lat_pm, long_pm) %>% 
  left_join(kids_amb_temp_data, by = "zip") %>% 
  select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)
kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
  mutate(sample_date = amb_pm_date) %>% 
  select(-lat_zip, -long_zip, -lat_pm, -long_pm)
kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  select(-lat_zip, -long_zip, -lat_temp, -long_temp)

@@

#write_rds(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.rds")
#write_csv(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.csv")
```

# Convert long data to medium: only keeping days with PM sampling, repeated by child_id
```{r}
exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")

@@

#write_rds(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.rds")
#write_csv(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.csv")
```

# Convert medium data to short: only keeping 1 observation per sampling period, repeated by child_id
```{r}
exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

@@

#write_rds(exposure_analysis_data_short, "Output/exposure_analysis_data_short.rds")
#write_csv(exposure_analysis_data_short, "Output/exposure_analysis_data_short.csv")
```

# Dataset with PM and SUMs matched/average over short interval
```{r}
pm_join <- pm_clean %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first_datetime_new,
         sample_datetime = round_date(sample_datetime, unit = "minute"),
         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
  ungroup() %>% 
  select(-home_id, -treatment) %>% 
  arrange(area, home, sample_datetime)

@@

exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")

@@

#write_rds(pm_sums_joined3, "Output/sums_rolling_pm_data.rds")
#write_csv(pm_sums_joined3, "Output/sums_rolling_pm_data.csv")
```

>>>>

---
title: "KidsAIR: merge exposure datasets"
author: "Ethan Walker"
date: "Started 24 March 2020, Updated 14 April 2020"
output: pdf_document
---

@@

# Load individual datasets
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

pm_clean <- 
  read_rds(paste0(output_path, "Output/pm_clean.rds"))
sums_clean <- 
  read_rds(paste0(output_path, "Output/sums_clean.rds"))

kids_linked_ids <- 
  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
demographics_clean <- 
  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
moisture_clean <- 
  read_rds(paste0(output_path, "Output/moisture_clean.rds"))
indoor_temp_rh_clean <- 
  read_rds(paste0(output_path, "Output/indoor_temp_rh_clean.rds"))
stove_use_clean <- 
  read_rds(paste0(output_path, "Output/stove_use_clean.rds"))
stove_grades_clean <- 
  read_rds(paste0(output_path, "Output/stove_grades_clean.rds"))
home_char_clean <- 
  read_rds(paste0(output_path, "Output/home_char_clean.rds"))
home_act_clean <- 
  read_rds(paste0(output_path, "Output/home_act_clean.rds"))
```

# Create new dataset with mean PM/SUMs; join with covariates
## Long data averaged over sampling date, including SUMs dates, repeated by child_id
```{r}
# Prep PM data to join with SUMs data by sampling day
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ",
           remove = FALSE) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
         pm_daily_obs = n()) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
         pm_total_observations, pm_mean_sampling_period, sampling_day, 
         pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
         sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
         pm_spike_2000, pm_spike_5000) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

#write_rds(pm_join, paste0(output_path, "Output/pm_join.rds"))
pm_join <- read_rds(paste0(output_path, "Output/pm_join.rds"))

@@

#write_rds(sums_join, paste0(output_path, "Output/sums_join.rds"))
sums_join <- read_rds(paste0(output_path, "Output/sums_join.rds"))

@@

#write_rds(pm_sums_joined1, paste0(output_path, "Output/pm_sums_joined1.rds"))
pm_sums_joined1 <- read_rds(paste0(output_path, "Output/pm_sums_joined1.rds"))

@@

#write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
#write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))

@@

#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
```

# Update long dataset -- REMEMBER TO UPDATE SHORT DATA BELOW
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date) %>% 
  distinct(zip, .keep_all = TRUE)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  distinct(zip, .keep_all = TRUE)

wmt_zipcodes <- read_csv(paste0(output_path, "Input/WMT/wmt_zipcodes.csv")) %>% 
  rename_all(tolower)

nn_zipcodes <- read_csv(paste0(output_path, "Input/NN/nn_zipcodes.csv")) %>% 
  rename_all(tolower)

ak_zipcodes <- read_csv(paste0(output_path, "Input/AK/ak_zipcodes.csv")) %>% 
  rename_all(tolower)

kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
  mutate(home = gsub(home, pattern = "_", replacement = ""),
         home = gsub(home, pattern = " ", replacement = "")) %>% 
  left_join(kids_amb_pm_data, by = "zip") %>% 
  select(home, zip, lat_pm, long_pm) %>% 
  left_join(kids_amb_temp_data, by = "zip") %>% 
  select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date) %>% 
  select(-lat_zip, -long_zip, -lat_pm, -long_pm)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  select(-lat_zip, -long_zip, -lat_temp, -long_temp)

@@

#write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
#write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
```

# Convert long data to medium: only keeping days with PM sampling, repeated by child_id
```{r}
exposure_analysis_data_long_new <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

@@

#write_rds(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
#write_csv(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.csv"))
```

# Convert medium data to short: only keeping 1 observation per sampling period, repeated by child_id
```{r}
exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

@@

#write_rds(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.rds"))
#write_csv(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.csv"))
```

# Dataset with PM and SUMs matched/average over short interval
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

pm_clean <- 
  read_rds(paste0(output_path, "Output/pm_clean.rds"))
sums_clean <- 
  read_rds(paste0(output_path, "Output/sums_clean.rds"))

pm_join <- pm_clean %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first_datetime_new,
         sample_datetime = round_date(sample_datetime, unit = "minute"),
         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
  ungroup() %>% 
  select(-home_id, -treatment) %>% 
  arrange(area, home, sample_datetime)

@@

exposure_analysis_data_long_new <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

@@

#write_rds(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.rds"))
#write_csv(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.csv"))
```
